The title

### 2024-12-10 05_Manuscript/manuscript.Rmd
updating with trackdown

```{r trackdown, eval = F}
library(trackdown)
# upload_file(
#   file = "manuscript.Rmd",
#   gfile = "gen_scale_manuscript_trackdown",
#   gpath = "gen_scale/",
#   hide_code = TRUE
# )

@@

## Material

## Procedure

```{r libraries}
library(rio)
library(dplyr)
set.seed(8943)
library(mice)
library(psych)

@@

@@

@@

@@

```{r data-combine-split}
combo <- bind_rows(sona_good, mturk_good)
combo_noout <- bind_rows(sona_noout, mturk_noout)

@@

efaDF <- combo %>% 
  filter(ResponseId %in% random_split) %>% 
  select(Q2_1:Q5_20)  %>% 
  select(-Q2_21, -Q4_20, -Q5_16)
cfaDF <- combo %>% 
  filter(!(ResponseId %in% random_split)) %>% 
  select(Q2_1:Q5_20)  %>% 
  select(-Q2_21, -Q4_20, -Q5_16)

@@

efaDF_noout <- combo_noout %>% 
  filter(ResponseId %in% random_split) %>% 
  select(Q2_1:Q5_20)  %>% 
  select(-Q2_21, -Q4_20, -Q5_16)
cfaDF_noout <- combo_noout %>% 
  filter(!(ResponseId %in% random_split)) %>% 
  select(Q2_1:Q5_20)  %>% 
  select(-Q2_21, -Q4_20, -Q5_16)

@@

```{r}
#efaDF <- read.csv("../03_Data/efaDF.csv")

@@

```{r}
item_text <- read.csv("../03_Data/item_text.csv")

@@

```{r}

@@

```{r}
EFA_fit <- fa(efaDF, #data
              nfactors = 5, #number of factors
              rotate = "oblimin", #rotation
              fm = "ml") #math

@@

- import the edited files and test those models for fit 
- theory is 5, so if the fit is similar, then go with five

--- 
test cfa 
test MGCFA --> use new package to include information 
---

**Model 1: Two-Factor**

These are all the items with loading \> 0.7

-   Factor One: Sense of Contribution

    -   Q2_16: I have taken part in activities or events which will
        impact the future.

    -   Q2_2: I have contributed to the world in a unique way.

    -   Q2_3: I have contributed to causes that will impact the future.

    -   Q5_9: I have contributed to things which will be meaningful.

    -   Q5_13: I will leave a lasting legacy.

    -   Q2_5: I am proud of my accomplishments.

    -   Q2_6: I am confident in the legacy I will leave when I die.

    -   Q4_1: Others would say I have impacted their lives.

<!-- -->

-   Factor Two: Spiritual Immortality

    -   Q4_19: I believe in an afterlife.

    -   Q2_17: My spirit or soul will live on after I physically die.

    -   Q2_12: I believe I will live on in the afterlife.

    -   Q5_2: Death is not the end.

    -   Q5_10: I believe in an immaterial soul or spirit.

    -   Q5_1: I am comforted by the thought of an afterlife.

    -   Q2_18: Death will not be the end of me.

**Model 2: Three-Factor**

-   Factor One:

    -   Q2_16: I have taken part in activities or events which will
        impact the future.

    -   Q2_3: I have contributed to causes that will impact the future.

    -   Q2_5: I am proud of my accomplishments.

    -   Q5_9: I have contributed to things which will be meaningful.

    -   Q4_1: Others would say I have impacted their lives.

    -   Q2_4: I feel I have taken part in something bigger than myself.

    -   Q2_2: I have contributed to the world in a unique way.

<!-- -->

-   Factor Two:

    -   Q

<!-- -->

-   Factor Three:

    -   Q

**Model 3: Five-Factor**

>>>>

```{r trackdown, eval = F}
library(trackdown)
source("~/GitHub/files/google_creds.R")
# upload_file(
#   file = "manuscript.Rmd",
#   gfile = "gen_scale_manuscript_trackdown",
#   gpath = "gen_scale/",
#   hide_code = TRUE
# )

@@

```{r libraries}
library(rio)
library(dplyr)
set.seed(8943)
library(mice)
library(psych)
library(visualizemi)
library(lavaan)

@@

```{r get-age}
sona <- sona %>% 
  mutate(Q13 = gsub("[a-z]", "", Q13)) %>% 
  mutate(Q13 = ifelse(
    Q13 == 1991, 2018-1991, Q13)
  ) %>% 
  mutate(Q13 = as.numeric(Q13))

mturk <- mturk %>% 
  mutate(Q14 = as.numeric(Q14))
```

Participants were recruited from two data sources. `r nrow(sona)`
undergraduate students taking introduction to psychology were recruited
for course credit at a large midwestern university. Undergraduate
participants were *M* = `r apa_num(mean(sona$Q13, na.rm = T))` (*SD* =
`r apa_num(sd(sona$Q13, na.rm = T))`) years old at the time of the
study. `r nrow(mturk)` participants were recruited from Amazon's
Mechanical Turk, and these participants were *M* =
`r apa_num(mean(mturk$Q14, na.rm = T))` (*SD* =
`r apa_num(sd(mturk$Q14, na.rm = T))`) years old. No other demographic
information was collected in the study.

## Material and Procedure

`r nrow(questions)-3` items were developed using previous scale
development on generativity [@drolet1990], along with theoretical factor
alignment to XXCITEXX. These items aligned with areas of biological,
physical, social, spiritual, and personal immortality. The survey was
implemented in Qualtrics. First, participants entered their age in years
for basic demographic information. The immortality items were separated
into three blocks of 20-21 questions. These three blocks of questions
were randomized across participants, and the items within each block
were also randomized. In each block, one attention check item was shown
(e.g., "Answer this question as somewhat agree."). At the end of the
study, participants were directed back to either SONA for undergraduate
course credit or Mechanical Turk for monetary
compensation[\^Unfortunately, the exact amount we paid participants is
lost to time. Given our ethics application, we believe it was 1-2 USD,
and the survey was completed in *M* =
`r apa_num(mean(mturk_good[, "Duration (in seconds)"])/60)` (*SD* =
`r apa_num(sd(mturk_good[, "Duration (in seconds)"])/60)` minutes.].

This study was not pre-registered. We conducted data analysis in two
phases. The data were first screened for inattentive and outlier
responses. The data were then randomly split into an exploratory and
confirmatory dataset. We developed the XX scale by using exploratory
factor analysis (EFA) with maximum likelihood estimation and an oblique
(i.e., oblimin) rotation [@preacher2003]. Analyses were performed with
the *psych* package [@revelle2020]. We used scree plots, parallel
analysis, and theory to determine the number of factors and compared
models based on fit indices to select the best model. We use the
following fit indices: Root Mean Squared Error of Approximation [RMSEA;
@steiger2016], Root Mean Squared Residual [RMSR; @jöreskog1996],
Comparative Fit Index [CFI; @bentler1990], and the Tucker-Lewis Index
[TLI; @tucker1973]. Values close to one are expected for goodness of fit
statistics [CFI, TLI; @bentler1990; @hu1999], while lower values close
to zero are considered better models for residual statistics [RMSEA,
RMSR; @steiger2016].

We then analyzed the best model from the exploratory stage using
confirmatory factor analysis (CFA) and multigroup CFA [@kline2016;
@brown2015] on the second confirmatory dataset. The models were analyzed
with *lavaan* [@rosseel2012] and *visualizemi* [@buchanan2024] *R*
packages. Multigroup CFA can be used to determine if the scale structure
can be generalized to subsamples within the data, and thus, allowing for
the same interpretation of total or subscale scores calculated from the
scale. First, we modeled the overall structure from the EFA results to
determine fit in a new sample. The RMSEA, CFI, and TLI were used to
determine model fit along with the Standardized Root Mean Squared
Residual [SRMR; @bentler1995]. The same interpretations were included
from the exploratory stage: high fit indices close to one for goodness
of fit statistics, and low fit indices close to zero for residual fit
statistics. If the overall model fit well, we then modeled the
subsamples of student populations and Mechanical Turk populations
separately. While fit is expected to potentially degrade due to smaller
sample sizes, we expect these models to converge to demonstrate that it
is appropriate to combine models for multigroup testing.

For the multigroup CFA, we used @brown2015's suggested terminology and
steps for analysis. The *configural* model includes both subsamples
combined together into one model controlling for group structure. This
model supports the "configuration" or structure of the scale is at least
similar within each group, but does not force each group to show similar
patterns of parameter estimates. Each subsequent step constrains a set
of parameters to be equal across groups to demonstrate invariance (i.e.,
equivalence). The configural model is then compared to the *metric*
model, which sets the factor loadings for each measured item to be equal
across groups. This model supports that each item is related to the
underlying latent variable the same way for each group. If this
assumption holds, the metric model is then compared to a *scalar* model,
which sets the item intercepts (i.e., item means) to be equal across
groups. Scalar model invariance suggests that each subsample shows the
same item means and factor loadings across groups. Finally, we compared
the scalar model to the *strict* model, which constrains item residuals
(i.e., errors, variance) to be the same across groups. Invariance in
model residuals suggests that the items have the same answer
range/spread across groups.

To assess if models are invariant, we used $\Delta$ CFI \<= .01
criterion to indicate that models were equivalent [@cheung2002]. If
models were not considered invariant, we investigated for the source of
the invariance by using partial invariance testing [@byrne1989;
@cheung1999]. Each parameter estimate is individually relaxed to
determine if that particular estimate leads to misfit in the model. For
example, if the model fails scalar invariance, each item mean would be
allowed to freely vary between groups (one at a time) to find the items
that show group differences. If only a few items show group differences,
then the model is considered "partially" invariant, and the impact on
total scores can be assessed. Partially invariant models indicate areas
for further investigation into group differences and ways that the final
scores of the model must be carefully interpreted. Previously, @nye2011
and @nye2019 proposed an effect size $d_{MACS}$ to help quantify the
size of the difference between groups in partially invariant models.
However, $d_{MACS}$ can generally only be used to understand the item
means and loadings together (due to the mathematical calculation),
regardless of the tyep of non-invariance. Therefore, we applied a new
technique proposed by @buchanan2024 to assess several potential effect
size components of the model: 1) overall model replication likelihood,
2) individual parameter replication likelihood, and 3) effect size for
individual parameters showing non-invariance.

We will use the $h_{nmi}$ effect size as the likelihood of replication
statistic for both the overall model and individual parameters. First,
the data is bootstrapped with replacement *n* times to create similar
samples to the original data. The grouping variable is then randomized
to create a comparison model. The proportion of times that invariance
(or non-invariance) is met is calculated for the bootstrapped and
randomized group labels. $h_{nmi}$ is Cohen's $h$ , which is the
standardized difference between two proportions for non-invariance (no
measurement invariance, but the statistic can also be calculated as
$h_{mi}$ for the difference in proportions for when the model showed
measurement invariance). We suggest $h_{nmi}$ because the interpretation
is a bit simpler: effect sizes close to zero indicate that the
bootstrapped data and randomized groups data show the same number of
non-measurement invariance findings, indicating that they are likely the
same result. The logic here is that if groups are truly equal in their
measurement on a scale, then the grouping variable should not affect the
results, thus, randomizing the grouping variable label does not affect
the results. As $h_{nmi}$ increases, we expect to find non-measurement
invariance between groups, as the bootstrapped data is consistently
different than randomized group labels. $h$ can be interpreted much like
$d$, but $d$ does not technically have upper or lower limits (i.e., they
are $\infty$), while $h$'s range is limited to $\pi$. Therefore, we will
report $h_{nmi_p}$ which is scaled to range from -1 to +1 for easier
interpretation. In addition to this effect size, we can calculate $d$
for the differences in parameters for bootstrapped versus random group
labels, thus, providing an estimate for the size of group differences on
partially invariant parameter estimates.

@@

### Undergraduate Data

Data were screened for accuracy, missing data, outliers, and multivariate parametric assumptions.

`r nrow(sona)`
`r nrow(sona_replaced)`
`r nrow(sona_good)`
`r nrow(sona_noout)`

@@

### Mechanical Turk Data

`r nrow(mturk)`
`r nrow(mturk_replaced)`
`r nrow(mturk_good)`
`r nrow(mturk_noout)`

@@

```{r data-combine-split}
combo <- bind_rows(sona_good %>% 
                     mutate(where = "Undergraduate"), 
                   mturk_good %>% 
                     mutate(where = "Mechanical Turk"))
combo_noout <- bind_rows(sona_noout %>% 
                     mutate(where = "Undergraduate"), 
                     mturk_noout %>% 
                     mutate(where = "Mechanical Turk"))

@@

efaDF <- combo %>% 
  filter(ResponseId %in% random_split) %>% 
  select(Q2_1:Q5_20)  %>% 
  select(-Q2_21, -Q4_20, -Q5_16)
cfaDF <- combo %>% 
  filter(!(ResponseId %in% random_split)) %>% 
  select(Q2_1:Q5_20, where)  %>% 
  select(-Q2_21, -Q4_20, -Q5_16)

@@

efaDF_noout <- combo_noout %>% 
  filter(ResponseId %in% random_split) %>% 
  select(Q2_1:Q5_20)  %>% 
  select(-Q2_21, -Q4_20, -Q5_16)
cfaDF_noout <- combo_noout %>% 
  filter(!(ResponseId %in% random_split)) %>% 
  select(Q2_1:Q5_20, where)  %>% 
  select(-Q2_21, -Q4_20, -Q5_16)

@@

```{r number-factors}
#efaDF <- read.csv("../03_Data/efaDF.csv")

@@

```{r two-factor}
item_text <- read.csv("../03_Data/item_text.csv")

@@

```{r three-factor}

@@

```{r five-factor}
EFA_fit <- fa(efaDF, #data
              nfactors = 5, #number of factors
              rotate = "oblimin", #rotation
              fm = "ml") #math

@@

```{r comparison-models}
efa_results <- import("results/final_questions.xlsx")
EFA_2factor <- fa(efaDF %>% 
                    select(efa_results %>% 
                           filter(!is.na(factor2)) %>% 
                           pull(question)),
                  nfactors = 2,
                  rotate = "oblimin", #rotation
                  fm = "ml") #math

EFA_3factor <- fa(efaDF %>% 
                    select(efa_results %>% 
                           filter(!is.na(factor3)) %>% 
                           pull(question)),
                  nfactors = 3,
                  rotate = "oblimin", #rotation
                  fm = "ml") #math

EFA_5factor <- fa(efaDF %>% 
                    select(efa_results %>% 
                           filter(!is.na(factor5)) %>% 
                           pull(question)),
                  nfactors = 5,
                  rotate = "oblimin", #rotation
                  fm = "ml") #math

EFA_2factor
EFA_3factor
EFA_5factor

new_5_columns <- as.data.frame(unclass(EFA_5factor$loadings)) %>% 
  mutate(question = rownames(.)) %>% 
  rowwise() %>%
  mutate(keep = sum(c_across(-question) >= .30)) %>% 
  ungroup() %>% 
  arrange(desc(keep)) %>% 
  filter(keep == 1) %>% 
  pull(question)

EFA_5factor <- fa(efaDF %>% 
                    select(all_of(new_5_columns)),
                  nfactors = 5,
                  rotate = "oblimin", #rotation
                  fm = "ml") #math

EFA_5factor

new_5_columns <- as.data.frame(unclass(EFA_5factor$loadings)) %>% 
  mutate(question = rownames(.)) %>% 
  rowwise() %>%
  mutate(keep = sum(c_across(-question) >= .30)) %>% 
  ungroup() %>% 
  arrange(desc(keep)) %>% 
  filter(keep == 1) %>% 
  pull(question)

EFA_5factor <- fa(efaDF %>% 
                    select(all_of(new_5_columns)),
                  nfactors = 5,
                  rotate = "oblimin", #rotation
                  fm = "ml") #math

EFA_5factor

# as.data.frame(unclass(EFA_5factor$loadings)) %>% 
#   mutate(question = rownames(.)) %>% 
#   rowwise() %>%
#   mutate(keep = sum(c_across(-question) >= .30)) %>% 
#   ungroup() %>% 
#   left_join(questions,
#             by = c("question" = "Question_Number")) %>% 
#   arrange(desc(ML1)) %>% 
#   export(., "results/5_factor_reduced.xlsx", row.names = F)
```

```{r}
# test if q2_1 changing to a different place changes results 
five_reduced <- import('results/5_factor_reduced.xlsx')
five.model <- ""

for (factor in unique(five_reduced$Factor)){
  questions <- five_reduced %>% 
    filter(Factor == factor) %>% 
    pull(question)

  five.model <- 
    paste0(five.model, "\n",
        factor, " =~ ", 
        paste(questions, collapse = " + "))
}

#cat(five.model)

model.results <- mgcfa(
  model = five.model,
  data = cfaDF,
  group = "where", 
  group.equal = c("loadings", "intercepts", "residuals"),
  meanstructure = TRUE
)

summary(model.results$model_overall, 
        standardized = T,
        fit.measures = T, 
        rsquare = T)

summary(model.results$group_models$model.Undergraduate, 
        standardized = T,
        fit.measures = T, 
        rsquare = T)

modificationindices(model.results$group_models$model.Undergraduate,
                    sort. = T)

summary(model.results$group_models$`model.Mechanical Turk`, 
        standardized = T,
        fit.measures = T, 
        rsquare = T)

modificationindices(model.results$group_models$`model.Mechanical Turk`, 
                    sort. = T)

summary(model.results$model_configural, 
        standardized = T,
        fit.measures = T, 
        rsquare = T)

model.results$model_fit

model.partial <- partial_mi(
  saved_model = model.results$invariance_models$model.intercepts,
  data = cfaDF,
  model = five.model,
  group = "where",
  group.equal = c("loadings", "intercepts"),
  partial_step = "intercepts")

model.partial$fit_table
```

```{r}
# distribution of each latent variable as a plot

saved_mi_plots <- plot_mi(data_coef = model.results.partial$model_coef,
                          # which model do you want to plot from model column
                          model_step = "intercepts", 
                          # name of observed item
                          item_name = "Q4_8", 
                          # LV limits to graph
                          x_limits = c(-2,2), 
                          # Y min and max in data
                          y_limits = c(min(cfaDF$Q4_8),
                                       max(cfaDF$Q4_8)), 
                          # what ci do you want
                          conf.level = .95, 
                          # what model results do you want
                          model_results = model.results.partial$invariance_models$model.intercepts, 
                          # which latent is the observed variable on
                          # important for cross-loaded variables
                          lv_name = "biological", 
                          # if you have more than two groups, which two do you want
                          plot_groups = NULL)

saved_mi_plots$complete

saved_mi_plots <- plot_mi(data_coef = model.results.partial$model_coef,
                          # which model do you want to plot from model column
                          model_step = "intercepts", 
                          # name of observed item
                          item_name = "Q2_4", 
                          # LV limits to graph
                          x_limits = c(-1,1), 
                          # Y min and max in data
                          y_limits = c(min(cfaDF$Q2_4),
                                       max(cfaDF$Q2_4)), 
                          # what ci do you want
                          conf.level = .95, 
                          # what model results do you want
                          model_results = model.results.partial$invariance_models$model.intercepts, 
                          # which latent is the observed variable on
                          # important for cross-loaded variables
                          lv_name = "social", 
                          # if you have more than two groups, which two do you want
                          plot_groups = NULL)

model.results.partial <- mgcfa(
  model = five.model,
  data = cfaDF,
  group = "where", 
  group.equal = c("loadings", "intercepts", "residuals"),
  meanstructure = TRUE,
  group.partial = "Q4_8 ~1"
)

# figure out how to update this 
# so that it will take the group.partial correctly 
model.partial.partial <- partial_mi(
  saved_model = model.results.partial$invariance_models$model.residuals,
  data = cfaDF,
  model = five.model,
  group = "where",
  group.equal = c("loadings", "intercepts", "residuals"),
  partial_step = "residuals")

model.partial.partial$fit_table
```

```{r boot-model, eval = F}
saved_boot_model <- bootstrap_model(
  # saved configural model to start at
  saved_configural = model.results$model_configural,
  # dataset for the analysis
  data = cfaDF,
  # model lavaan syntax
  model = five.model,
  # group variable in the dataset
  group = "where", 
  # number of bootstraps
  # this is set to a low number to compile quickly for cran
  nboot = 1000,
  # name of the fit measure you want to use, make sure it's lavaan
  invariance_index = "cfi",
  # rule for the difference in fit indices
  invariance_rule = .01,
  # what order of steps do you want to test? 
  group.equal = c("loadings", "intercepts", "residuals")
)

saveRDS(saved_boot_model, "../03_Data/saved_boot_model.rds")
```

```{r boot-partial, eval = F}
saved_boot_partial <- bootstrap_partial(
  # the model you want to test 
  # use the model before your invariant one 
  # similar set up to partial_mi
  saved_model = model.results$invariance_models$model.loadings,
  # the dataframe
  data = cfaDF,
  # the original model syntax
  model = five.model,
  # the grouping variable column
  group = "where",
  # run more, but this package vignette needs to knit fast
  nboot = 1000,
  # what index are you using for invariance?
  # match this to lavaan's name under fitmeasures()
  invariance_index = "cfi",
  # what rule are you using?
  invariance_rule = .01,
  # what are we comparing against? 
  invariance_compare = unname(fitmeasures(model.results$invariance_models$model.loadings, "cfi")),
  # which step you want to estimate effect size for
  partial_step = c("intercepts"), 
  # which parameters you want to hold constrained
  group.equal = c("loadings", "intercepts")
  )
saveRDS(saved_boot_partial, "../03_Data/saved_boot_partial.rds")
```
